Evolutionary stretegies

### 2016-01-23 evolutionaryComputation/summery.tex
Reduced the text size significantly and continued steamrolling trough the slides. Also added my proper name to the text instead of the stackoverflow thing.

\documentclass[10pt,landscape]{article}
\usepackage{multicol}
\usepackage{calc}
\usepackage{ifthen}
\usepackage[landscape]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage{color,graphicx,overpic}
\usepackage{hyperref}

\pdfinfo{
  /Title (example.pdf)
  /Creator (TeX)
  /Producer (pdfTeX 1.40.0)
  /Author (Seamus)
  /Subject (Example)
  /Keywords (pdflatex, latex,pdftex,tex)}

@@

>>>>

\documentclass[8pt,landscape]{extarticle}
\usepackage{multicol}
\usepackage{calc}
\usepackage{ifthen}
\usepackage[landscape]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage{color,graphicx,overpic}
\usepackage{hyperref}

\pdfinfo{
  /Title (EC cheatsheet)
  /Creator (TeX)
  /Producer (pdfTeX 1.40.0)
  /Author (Jappie Klooster)
  /Subject (Evulotionary computing)
  /Keywords (Evolutionary, Computing, Gentic, Algorithm, Heuristics, Strategy)}

@@

\section{Multi objective EA's}
\subsection{Pareto}
\emph{Dominance} A solution $x^0$ is said to (Paretor) \emph{dominate} a
solution $x^1$ (denoted $x^0\succ x^1$) iff
$(\forall i \in \{0,1,\dots,m-1\}:f_i(x^0)\leq f_i(x^1))\wedge
 (\exists i \in \{0,1,\dots,m-1\}:f_i(x^0)< f_i(x^1))$.
A solution is pareto optimal iff $\neg\exists x^1:x^1\succ x^0$.
The parote optimall set: $P_s=\{x^0|\neg\exists x^1:x^1\succ x^0\}$.
The paroto optimal front is the $P_f$ of all objective function values
correspoinding to the solutions in $P_s$: $P_F=\{f(x)=(f_0(x),f_1(x),\dots,
	f_{m-1}(x))|x\in P_s$
\subsection{Weighted aggregation}
The graphs with the linerider lines. It doesn't work with bowl
like objectives because only the edges are found.
\subsection{Evolutionary Multi-objective Optimization}
Desired result: representative subset of $P_s$.
\emph{Domination count}
For each solution count how many times
its dominatad. Then use one-dimensional selection (truncate or tournament).
\emph{Domination rank} solutions of rank 0 have domination count 0, solutions
of rank i are solutions of rank 0 when solutions of lower ranks are disregarded.

\subsection{Elitism}
What if number of rank 0 solutions is bigger than the next generation?
Must diregard some of the rank 0 solutions. The space dominated by
elitist solutions changes, this can cause the elitist solutions/front to
become \emph{worse}. This isn't true eltism!
For this reason an elitistarchive is often used. Current best solutions
are stored in the archive if they dominate at least one solution in the
archive or if they are not dominated by any solution in the archive.
Dominated solutions are removed from the archive. It can grow exponentially
large. Pruning the archive causes the same problems again. One solution:
discretize objective space, store at most one solution per
discretization box. Typically for better convergence also select
solutions from archive.
\subsection{Performance indicators}
Performance indicator is a function: Input approximation set, output: single
value representing a certaing goodnes. However no single or combination of
performance indiciators taht compare approximation sets can always tell wether
one approximation set is better than another.

\section{Model-based EAs: Est. of dist. algorithms}
Model based evolutionary algorthms: Population based stochastic search
algorithms. Exploitation: selection, Exploration: Learn model from selected
solutions, generate new solutions from the model (\&pop).
\subsection{Univariate PMBGA}
No statistical interaction between variables considered.
Model: probability vector $[p_1,\dots,P_l]$($l$:string length).
$p_i$: probability of value 1 at string position $i$.
$p(x)=\Pi^l_{i=1}p(x_i) \quad$ ($p(x_i)$: univariate marginal distribution).
Learn modal: count proportions of 1 in selected population. Sample model:
Generate new solutions with specified probabilities.

%TODO: lack of space remove variants
\subsubsection{Variants}
PBIL: Prob. vector incrementally updated over successive generations.
UMDA: No incremental updates. Compact GA: Models steady-state GA with
tournament selection. DEUM: Uses markov random field modeling.

\subsection{Bivariate PMBGA}
Need more than just probabilities of bit values, model pairwise interactions:
conditional probablities.

\subsubsection{Ntropy}
Random variable $X$ with probability istrubtion function $p(X)$. Entropy
$H(X$ is a measure of unceartainty about a random variable $X$:
$H(X) =\sum_{x\in X}-p(x)\mbox{log }p(x)$. Conditionaly entropy $H(Y|X)$ is
a measure of uncertainty remianing about $Y$ after $X$ is known (what $X$
does not say about $Y$):
$H(Y|X) =\sum\limits_{x\in X}\sum\limits_{y\in Y}p(x,y)\mbox{log }
\frac{p(x)}{p(x,y)}$. The mutual information $I(X,Y)$ of two random variables
is a measure of the variables' mutual dependence. Mutual information determines
how similiar the joint distribution $P(X,Y)$ is to the products of factored
marginal distribution $p(X)p(Y)$:
$I(Y,X) =\sum\limits_{x\in X}\sum\limits_{y\in Y}p(x,y)\mbox{log }
\frac{p(x,y)}{p(x)p(y)}$. Mutual information in relation to Ntropy:
$I(X,Y) = H(Y)-H(Y|X)=H(X)+H(Y)-H(X,Y)$.
